---
title: "Particle Collision and Handwritten Digits Data Analysis"
author: "Arin Vansomphone"
date: "`r Sys.Date()`"
output: html_document
---
# Loading the packages
```{r}
library(tidyverse)
library(readr)
```

# Importing the data
### Particle data
```{r warning=FALSE}
particledata <- read_csv("samples_test_175_data.csv")
particledata <- particledata %>%
  select(-...1)

particledata_long <- particledata %>%
  select(-`550 samples`, -`600 samples`, -`650 samples`, -`700 samples`) %>%
  rename(`50` = `50 samples`,
         `100` = `100 samples`,
         `150` = `150 samples`,
         `200` = `200 samples`,
         `250` = `250 samples`,
         `300` = `300 samples`,
         `350` = `350 samples`,
         `400` = `400 samples`,
         `450` = `450 samples`,
         `500` = `500 samples`,) %>%
  pivot_longer(names_to = "samplesize",
               values_to = "percentaccuracy",
               '50' : '500')
particledata_long$samplesize <- as.double(particledata_long$samplesize)
particledata_long
```

### Handwritten data
```{r warning=FALSE}
handwrittendata <- read_csv("handwrittendigits175testdata.csv")
handwrittendata <- handwrittendata %>%
  select(-...1)

handwrittendata_long <- handwrittendata %>%
  select(-`550 samples`, -`600 samples`, -`650 samples`, -`700 samples`) %>%
  rename(`50` = `50 samples`,
         `100` = `100 samples`,
         `150` = `150 samples`,
         `200` = `200 samples`,
         `250` = `250 samples`,
         `300` = `300 samples`,
         `350` = `350 samples`,
         `400` = `400 samples`,
         `450` = `450 samples`,
         `500` = `500 samples`,) %>%
  pivot_longer(names_to = "samplesize",
               values_to = "percentaccuracy",
               '50' : '500')
handwrittendata_long$samplesize <- as.double(handwrittendata_long$samplesize)
handwrittendata_long
```

# Finding the summary statistics
### Particle data
```{r}
particledatastats <- particledata_long %>%
  group_by(samplesize) %>%
  summarize(average = mean(percentaccuracy),
            sd = sd(percentaccuracy),
            min = min(percentaccuracy),
            Q1 = quantile(percentaccuracy, probs = 0.25),
            median = median(percentaccuracy),
            Q3 = quantile(percentaccuracy, probs = 0.75),
            max = max(percentaccuracy),
            IQR = Q3 - Q1) %>%
  arrange(samplesize)
particledatastats
```

### Handwritten data
```{r}
handwrittendatastats <- handwrittendata_long %>%
  group_by(samplesize) %>%
  summarize(average = mean(percentaccuracy),
            sd = sd(percentaccuracy),
            min = min(percentaccuracy),
            Q1 = quantile(percentaccuracy, probs = 0.25),
            median = median(percentaccuracy),
            Q3 = quantile(percentaccuracy, probs = 0.75),
            max = max(percentaccuracy),
            IQR = Q3 - Q1) %>%
  arrange(samplesize)
handwrittendatastats
```

# Plotting the data
### Particle data
```{r}
particledata_long %>%
  ggplot(aes(x = samplesize, y = percentaccuracy)) +
  geom_point() + 
  labs(x = "Sample Size",
       y = "Accuracy (%) for 175 test samples")
```

### Handwritten data
```{r}
handwrittendata_long %>%
  ggplot(aes(x = samplesize, y = percentaccuracy)) +
  geom_point() + 
  labs(x = "Sample Size",
       y = "Accuracy (%) for 175 test samples")
```

# Statistical tests
### Particle data
Line of best fit
```{r}
x = particledata_long$samplesize
y = particledata_long$percentaccuracy

plot(x, y, col = 'red', cex = 1.2)
abline(lm(y ~ x), col = 'blue')
```
```{r}
summary(lm(y ~ x))$coefficients
```
P value of roughly 0.148 is greater than $\alpha$ = 0.05, so we fail to reject the null hypothesis. There is insufficient evidence that greater sample sizes increase the model's accuracy. 

R-squared value
```{r}
particleml = lm(y ~ x, data = particledata_long)
summary(particleml)$r.squared
```

### Handwritten data
Line of best fit
```{r}
a = handwrittendata_long$samplesize
b = handwrittendata_long$percentaccuracy

plot(a, b, col = 'red', cex = 1.2)
abline(lm(b ~ a), col = 'blue')
```
```{r}
summary(lm(b ~ a))$coefficients
```
P value of approximately 0 is less than $\alpha$ = 0.05, so we reject the null hypothesis. There is strong evidence that increasing the sample size with the handwritten digits data increases the model's accuracy.

R-squared value
```{r}
handwrittenml = lm(b ~ a, data = handwrittendata_long)
summary(handwrittenml)$r.squared
```

# Particle data graph of standard deviation
```{r}
particledatastats %>%
  ggplot(aes(x = samplesize, y = sd)) + 
  geom_point() +
  labs(x = "Sample Size",
       y = "Standard Deviation of Accuracy (%) for 175 Test Samples") +
  geom_smooth(se = FALSE, method = 'lm')
```

### Statistical test for standard deviation
```{r}
c = particledatastats$sd
d = particledatastats$samplesize
summary(lm(c ~ d))$coefficients
```
P value of approximately 0 is less than $\alpha$ = 0.05, so we reject the null hypothesis. There is strong evidence that increasing the sample size with the particle data decreases the standard deviation, and thus spread, of the model's accuracy.

# Boxplot graphs for accuracy
### Particle data
```{r}
ggplot(particledata_long, aes(x = samplesize, y = percentaccuracy)) +
  geom_boxplot(aes(group = samplesize)) +
  labs(x = "Sample Size",
       y = "Accuracy (%) for 175 Test Samples")
```

### Handwritten data
```{r}
ggplot(handwrittendata_long, aes(x = samplesize, y = percentaccuracy)) +
  geom_boxplot(aes(group = samplesize)) +
  labs(x = "Sample Size",
       y = "Accuracy (%) for 175 Test Samples")
```

