---
title: "Particle Collision and Handwritten Digits Data Analysis"
author: "Arin Vansomphone"
date: "`r Sys.Date()`"
output: html_document
---
# Loading the Packages
```{r}
library(tidyverse)
library(readr)
```
# Sample Size Analysis
## Importing and Tidying the Data
```{r warning=FALSE}
particledata <- read_csv("Particle Physics Sample Data - Sheet1.csv")

particledata_long <- particledata %>%
  pivot_longer(names_to = "samplesize",
               values_to = "percentaccuracy",
               '400' : '3400')
particledata_long$samplesize <- as.double(particledata_long$samplesize)
particledata_long
```

## Finding the Summary Statistics
```{r}
particledatastats <- particledata_long %>%
  group_by(samplesize) %>%
  summarize(average = mean(percentaccuracy),
            sd = sd(percentaccuracy),
            min = min(percentaccuracy),
            Q1 = quantile(percentaccuracy, probs = 0.25),
            median = median(percentaccuracy),
            Q3 = quantile(percentaccuracy, probs = 0.75),
            max = max(percentaccuracy),
            IQR = Q3 - Q1) %>%
  arrange(samplesize)
particledatastats
```

## Plotting the Data
```{r}
particledata_long %>%
  ggplot(aes(x = samplesize, y = percentaccuracy)) +
  geom_point() + 
  labs(x = "Training Set Size",
       y = "Accuracy (%) for 100 test samples")
```

## Linear Regression Tests
### Weighted least squares
```{r}
x = particledata_long$samplesize
y = particledata_long$percentaccuracy

model <- lm(y ~ x)
```

```{r}
# test for heteroscedasticity
plot(fitted(model), resid(model), xlab = 'Fitted Values', ylab = 'Residual Values')

abline(0,0)
```

```{r}
# breusch-pagan test
library(lmtest)

bptest(model)
```

```{r}
wt <- 1 / lm(abs(model$residuals) ~ model$fitted.values)$fitted.values^2

wls_model <- lm(y ~ x, weights = wt)

summary(wls_model)
```

### Log transformation
```{r}
particledata_long$log_accuracy <- log(particledata_long$percentaccuracy)
particledata_long %>%
  ggplot(aes(x = samplesize, y = log_accuracy)) +
  geom_point() + 
  labs(x = "Sample Size",
       y = "Accuracy (%) for 175 test samples")
```
```{r}
z <- particledata_long$log_accuracy

log_model <- lm(z ~ x)

summary(log_model)
```

### Least squares
```{r}
particledata_long %>%
  ggplot(aes(x = samplesize, y = percentaccuracy)) +
  geom_point() + 
  labs(x = "Training Set Size",
       y = "Accuracy (%) for 100 test samples") +
  geom_smooth(se = FALSE, method = "lm") +
  geom_line(aes(y=20), color = "red")
```
```{r}
x = particledata_long$samplesize
y = particledata_long$percentaccuracy

model <- lm(y ~ x)

summary(model)
```

## Model accuracy vs. random guessing
```{r}
particledata500_long <- particledata_long %>%
  filter(samplesize == 3400)

t.test(particledata500_long$percentaccuracy, mu = 0.20, alternative = 'greater')
```

## Accuracy Standard Deviation vs. Training Sample Size
```{r}
particledatastats %>%
  ggplot(aes(x = samplesize, y = sd)) + 
  geom_point() +
  labs(x = "Sample Size",
       y = "Standard Deviation of Accuracy (%) for 100 Test Samples") +
  geom_smooth(se = FALSE, method = 'lm')
```

### Statistical test for standard deviation
```{r}
c = particledatastats$sd
d = particledatastats$samplesize
summary(lm(c ~ d))$coefficients
```

## Boxplot graph
```{r}
ggplot(particledata_long, aes(x = samplesize, y = percentaccuracy)) +
  geom_boxplot(aes(group = samplesize)) +
  labs(x = "Training Set Size",
       y = "Accuracy (%) for 100 Test Samples")
```

# Epoch Data
## Importing and Tidying the Data
```{r warning=FALSE}
epochdata <- read_csv("Particle Physics Epoch Data - Sheet1.csv")

epochdata_long <- epochdata %>%
  pivot_longer(names_to = "epochs",
               values_to = "percentaccuracy",
               '3' : '500')

epochdata_long$epochs <- as.double(epochdata_long$epochs)
epochdata_long <- epochdata_long %>%
  na.omit()
```

## Finding the Summary Statistics
```{r}
epochdatastats <- epochdata_long %>%
  group_by(epochs) %>%
  summarize(average = mean(percentaccuracy),
            sd = sd(percentaccuracy),
            min = min(percentaccuracy),
            Q1 = quantile(percentaccuracy, probs = 0.25),
            median = median(percentaccuracy),
            Q3 = quantile(percentaccuracy, probs = 0.75),
            max = max(percentaccuracy),
            IQR = Q3 - Q1) %>%
  arrange(epochs)
epochdatastats
```

## Plotting the Data
```{r}
epochdata_long %>%
  filter(epochs == c(3, 4, 5, 6, 7)) %>%
  ggplot(aes(x = epochs, y = percentaccuracy)) +
  geom_point() + 
  labs(x = "Epochs",
       y = "Accuracy (%) for 100 test samples")
```

## Least squares
```{r}
epochdata_long %>%
  filter(epochs == c(3, 4, 5, 6, 7)) %>%
  ggplot(aes(x = epochs, y = percentaccuracy)) +
  geom_point() + 
  labs(x = "Epochs",
       y = "Accuracy (%) for 100 test samples") +
  geom_smooth(se = FALSE, method = "lm") +
  geom_line(aes(y=20), color = "red")
```

```{r}
epochdata_long_small <- epochdata_long %>%
  filter(epochs == c(3, 4, 5, 6, 7))

e = epochdata_long_small$epochs
f = epochdata_long_small$percentaccuracy

epochmodel <- lm(f ~ e)

summary(epochmodel)
```

## Accuracy Standard Deviation vs. Training Sample Size
```{r}
epochdatastats %>%
  filter(epochs == c(3, 4, 5, 6, 7)) %>%
  ggplot(aes(x = epochs, y = sd)) + 
  geom_point() +
  labs(x = "Epochs",
       y = "Standard Deviation of Accuracy (%) for 100 Test Samples") +
  geom_smooth(se = FALSE, method = 'lm')
```

### Statistical test for standard deviation
```{r}
epochdatastats_small <- epochdatastats %>%
  filter(epochs == c(3, 4, 5, 6, 7)) 

g = epochdatastats_small$sd
h = epochdatastats_small$epochs
summary(lm(h ~ g))$coefficients
```

# Binary Classification Data
## Importing and Tidying the Data
```{r warning=FALSE}
binarydata <- read_csv("Binary Test Data - Sheet1.csv")

binarydata_long <- binarydata %>%
  pivot_longer(names_to = "classification",
               values_to = "percentaccuracy",
               'Muon' : 'Proton')
```

## Finding statistics
```{r}
binarydatastats <- binarydata_long %>%
  group_by(classification) %>%
  summarize(average = mean(percentaccuracy),
            sd = sd(percentaccuracy),
            min = min(percentaccuracy),
            Q1 = quantile(percentaccuracy, probs = 0.25),
            median = median(percentaccuracy),
            Q3 = quantile(percentaccuracy, probs = 0.75),
            max = max(percentaccuracy),
            IQR = Q3 - Q1) %>%
  arrange(classification)
binarydatastats
```

## Plotting the data
```{r}
binarydata_long %>%
  ggplot(aes(x = classification, y = percentaccuracy)) +
  geom_point() + 
  labs(x = "Classification",
       y = "Accuracy (%) for 100 test samples")
```

## Boxplot graph
```{r}
ggplot(binarydata_long, aes(x = classification, y = percentaccuracy)) +
  geom_boxplot(aes(group = classification)) +
  labs(x = "Classification",
       y = "Accuracy (%) for 100 Test Samples") +
  scale_y_continuous(limits = c(0,100))
```

## ANOVA Test
```{r}
binary.aov <- aov(percentaccuracy ~ classification, data = binarydata_long)

summary(binary.aov)
```
```{r}
TukeyHSD(binary.aov)
```

